No Massive Companion to the low-frequency, coherent radio-emitting M~Dwarf GJ\,1151

### 2019-06-21 paper/ms.tex
Joe's additions, comments, and edits

\title{No Massive Companion to the M~Dwarf Radio Source GJ\,1151}

@@

\author[0000-0002-0872-181X]{Harish Vedantham}
\affiliation{ASTRON, Netherlands Institute for Radio Astronomy, Oude Hoogeveensedijk 4, Dwingeloo, 7991 PD, The Netherlands}

@@

%% Mark off the abstract in the ``abstract'' environment. 
\begin{abstract}
The recent detection of circularly polarized, variable low-frequency radio emission from the M4.5~dwarf GJ\,1151 has been interpreted as a Jupiter-like electron cyclotron maser instability arising from star-planet interaction. The existence or parameters of the proposed planets have not been determined. In order to search for any planets or binary companions, we have obtained new HARPS-N observations

@@

Exoplanet science has flourished over the last three decades to be one of the major topics in optical astronomy. The number of known planets has doubled nearly every two years since 1995 \citep{Mamajek2016} and this accelerating rate of discovery is projected to continue for at least the next decade if current and upcoming space-based surveys deliver their expected results. However despite extensive searches \citep{2000ApJ...545.1058B,2013A&A...552A..65L,Lynch18} neither exoplanets nor their host stars have been detected at radio frequencies, as the quiescent emission of such systems has been too faint for current telescopes. But as the Square Kilometre Array \citep[SKA;][]{2009IEEEP..97.1482D} and its precursors, in particular LOFAR \citep[the LOw-Frequency ARray:][]{lofar} come online, the orders-of-magnitude increase in sensitivity and survey speed they provide makes the detection of nearby stars and planets a realistic prospect \citep{pope19}.

@@

Rather than explicitly searching for radio emission from known exoplanet hosts, \citet{Callingham_2019} cross-matched sources identified by the LOFAR Two-meter Sky Survey \citep[LoTSS:][]{lotss} with nearby stellar sources found by \gaia, finding the great majority of matches to be chance associations. But by restricting the survey to variable and circularly-polarized emission, the rate of chance associations with background radio galaxies is dramatically reduced. Based on this restricted cross-match, Vedantham et al. (submitted) have detected the M4.5 dwarf GJ\,1151 at low radio frequencies with LOFAR, finding a high degree of circular polarization and low X-ray luminosity consistent with a possible star-planet magnetic interaction. While M~dwarfs are known to flare at low frequencies \citep[e.g.][]{Villadsen19}, this emission shows a stead on/off duty cycle like that of electron cyclotron maser instability (ECMI) emission from Jupiter.

In this Letter we analyze new HARPS-N \citep[High Accuracy Radial velocity Planet Searcher:][]{harpsn} observations of GJ\,1151 in order to search for radial velocity signals of the proposed planets. We do not detect any planets, but place strong upper limits of a few Earth masses on the $M\sin{i}$ of any possible companions, ruling out any short-period massive objects or close binary companion.

@@

We obtained 20 epochs of observations of GJ\,1151 with HARPS-N from 2018-12-20 to 2019-02-27, as a Director's Discretionary Time program. While RVs were extracted from these using the standard HARPS pipeline, its performance on this M4.5 dwarf was very poor, resulting in a spurious RV scatter of several km/s. We therefore reprocessed these data using \textit{wobble} \citep{wobble}, a data-driven package which simultaneously non-parametrically constructs a stellar spectral template and telluric spectral components and uses these, rather than model spectral masks, to extract radial velocities.

>>>>

\title{No Massive Companion to the low-frequency, coherent radio-emitting M~Dwarf GJ\,1151}

@@

\author[0000-0002-0872-181X]{Harish K. Vedantham}
\affiliation{ASTRON, Netherlands Institute for Radio Astronomy, Oude Hoogeveensedijk 4, Dwingeloo, 7991 PD, The Netherlands}

@@

%% Mark off the abstract in the ``abstract'' environment. 
\begin{abstract}
The recent detection of circularly polarized, variable low-frequency radio emission from the M4.5~dwarf GJ\,1151 has been interpreted as a Jupiter-like electron cyclotron maser instability arising from a star-planet interaction. The existence or parameters of the proposed planets have not been determined. In order to search for any planets or binary companions, we have obtained new HARPS-N observations

@@

Exoplanet science has flourished over the last three decades. The number of known planets has doubled nearly every two years since 1995 \citep{Mamajek2016} and this accelerating rate of discovery is projected to continue for at least the next decade if current and upcoming space-based surveys deliver their expected results. However, despite extensive searches \citep[e.g.][]{2000ApJ...545.1058B,2013A&A...552A..65L,Lynch18}, neither exoplanets nor their host stars have been detected at radio frequencies, as the quiescent emission of such systems has been too faint for current telescopes \textcolor{red}{JRC - Not clear what you mean by quiescent here? E.g. Is ECMI quiescent emission?}. But as the Square Kilometre Array \citep[SKA;][]{2009IEEEP..97.1482D} and its precursors, in particular LOFAR \citep[the LOw-Frequency ARray:][]{lofar} come online \textcolor{red}{JRC - LOFAR has been operating for a decade, is not ``coming online''. I would drop the mention of SKA and just say LOFAR, an SKA-precursor, is unparalleled sensitivity at 150 MHz  where detections of such interaction are now possible and cite Zarka et al. (his scaling work)}, the orders-of-magnitude increase in sensitivity and survey speed they provide makes the detection of nearby stars and planets a realistic prospect \citep{pope19}.

@@

Rather than explicitly searching for radio emission from known exoplanet hosts, \citet{Callingham_2019} cross-matched sources identified by the LOFAR Two-meter Sky Survey \citep[LoTSS:][]{lotss} with nearby stellar sources found by \gaia, finding the great majority of matches to be chance associations. But by restricting the survey to variable and circularly-polarized emission, the rate of chance associations with background radio galaxies is dramatically reduced. Based on this restricted cross-match, Vedantham et al. (submitted) have detected the quiescent M4.5 dwarf GJ\,1151 at low radio frequencies with LOFAR that has properties that suggest the low-frequency radio emission is driven by a star-planet magnetic interaction. While M~dwarfs are known to flare at low frequencies \citep[e.g.][]{Villadsen19}, this emission lasts over 8\,h and is $> 70\%$ circularly polarised. Such emission can be generated by the electron cyclotron maser instability (ECMI) through the interaction of the star with a short ($\lesssim$ 10\,day) period planet or a close stellar companion as seen in UV Ceti-like stars \citep{2017ApJ...836L..30L}.

Since the radio emission implies a potential planetary or stellar companion, in this Letter we analyze take and analyse HARPS-N \citep[High Accuracy Radial velocity Planet Searcher:][]{harpsn} observations of GJ\,1151 in order to search for radial velocity signals of the proposed companions. We do not detect any planets, but place strong upper limits of a few Earth masses on the $M\sin{i}$ of any possible companions, ruling out any short-period massive objects or close binary companion.

@@

We obtained 20 epochs of observations of GJ\,1151 with HARPS-N from 2018-12-20 to 2019-02-27, as a Director's Discretionary Time program (Program ID: A38DDT2; PI: Callingham). While RVs were extracted from these using the standard HARPS pipeline, its performance on this M4.5 dwarf was very poor, resulting in a spurious RV scatter of several km/s. We therefore reprocessed these data using \textit{wobble} \citep{wobble}, a data-driven package which simultaneously non-parametrically constructs a stellar spectral template and telluric spectral components and uses these, rather than model spectral masks, to extract radial velocities.
